The onset of hydrodynamically driven waving of marine grass in flow

### 2014-09-26 Paper/grass.tex
Changes made by Mahesh
1) Added acknowledgments for myself.
2) Made few minor typo corrections (mostly singular/plural forms).
3) Changed symbol for Reynolds from R to Re
4) Replaced "figure" with "fig." for figure references and added "eq."
before all equation references.
5) Changed Order of magnitude symbol to math style rather than text.

\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{\text{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami }}
% \newcommand{\shreyas}[1]{{\bf (#1)}}
\newcommand{\shreyas}[1]{}
\definecolor{tableShade}{gray}{0.8}

\begin{document}
\title{The onset of hydrodynamically driven waving of marine grass in flow}
\author{Ravi Singh}
\affiliation{Brown University, Providence RI 02912 USA}
\author{L. Mahadevan}
\affiliation{Harvard University, Cambridge MA 02138 USA}
\author{M. M. Bandi\thanks{work performed while visiting Brown University}}
\affiliation{Collective Interactions Unit, OIST Graduate University, 1919-1 Tancha, Onna-son, Okinawa, Japan 904-0495}
\author{Amala Mahadevan}
\affiliation{Woods Hole Institute of Oceanography, Woods Hole MA USA}
\author{Shreyas Mandre}
\affiliation{Brown University, Providence RI 02912 USA}

@@

We develop a self-consistent model of the flow past rigid vegetation, where the vegetation is represented by a continuum drag field.
To explain the onset of spontaneous waving, we perform a linear stability analysis of the steady flow resulting from this model.
We show that a threshold flow condition for the spontaneous waving of marine grass arises naturally in this framework due to the drag exerted by the vegetation on the flow. 
Beyond the threshold, we find two unstable modes; first is localized in a boundary layer near the canopy top, and second which spans the whole water column.
The first mode shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it. 
The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations. 

@@

While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspect of existing theory remain unsatisfactory. 
(i) While the free shear layer is established due to the vegetation drag, linear stability analysis\cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
(ii) Classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold 

@@

Here we present a self-consistent mathematical model that account correctly for these effects 

@@

The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
The flexibility of the grass blades, on the other hand, serves merely to aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 

@@

Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, 

@@

where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 

@@

where $\rho$ the fluid density, $\mathbf{u}$ the velocity, 

@@

The Reynolds number of the flow based on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by pressure and modeled as 

@@

normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 

@@

In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.

@@

To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 

@@

A comparison of the steady flow profile from the solution of \eqref{base_equ} with experimental measurements is shown in figure~\ref{basicflow}.
Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient. 

@@

\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf\cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 

@@

This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami. 

@@

This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of figure \ref{basicflow}.

@@

To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in \eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)

@@

With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 

@@

 parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 

@@

of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 

@@

We solve this eigenvalue problem and a threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), emerges from the 
solution of \eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
The threshold Reynolds number increases with the grass density, indicating a competition between the 
destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of ad-hoc
assumptions for the mean velocity profile\cite{Raupach96,Delangre06}.

@@

Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in figure \ref{Re_vs_delta}b. 
The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
%Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.

\begin{table*}
\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
\renewcommand{\arraystretch}{1.4}
 \begin{tabular}{l|c|c|c}
			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
 Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
 Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
 Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
 Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
 Linearized dynamics	& $\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
 Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
 Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
 Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no
 \end{tabular}
 \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
 \label{tab:comparison}
\end{table*}
To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see figure \ref{Re_vs_delta}). 
To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
The lowest $\Rey$ on the neutral curve sets the threshold. 
We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the most unstable mode to jump.

The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold. 
This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$. 
The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$) leading to a simplification of \eqref{Orr-somerfield} as
\begin{align}
(D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \Rey \sigma (D^2-k^2)\phi,
\label{eqn:mode1asymp}
\end{align} 
in a region of thickness O($\delta$) near $y=\hg$.
Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}). 
This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.

On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
This estimate simplifies \eqref{Orr-somerfield} to 
\begin{align}
\begin{split}
\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,  \quad &\text{ for $y<\hg$}  \\
\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, \quad &\text{ for $y>\hg$}.
\end{split}
\label{eqn:mode2asymp}
\end{align}
\begin{figure}
\includegraphics[]{Asymptotic_noshear}
\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. 
Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 1 shapes appear self-similar in shape as $\delta\to 0$, 
and are compared to each other on a rescaled coordinate in the inset. 
Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. 
The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
\label{Asymptotic_mode}
\end{figure}
Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in figure \ref{Asymptotic_mode}. 
We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role. The only remaining parameter $\Rey/\Ndg$ 
sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
%On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.

@@

The appearance of vegetation drag parameter in the dominant balances \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and the threshold criteria demonstrates its role in setting the threshold.

@@

Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami}\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}. 
Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades. We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant. In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case. 

@@

The deviation of our model predictions from the observed may be attributed to the various simplification we have made in our model. 
In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. 
The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}. 

@@

\bibliography{Grass}{}
\bibliographystyle{plain}
\end{document}

>>>>

\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{\text{Re}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami }}
% \newcommand{\shreyas}[1]{{\bf (#1)}}
\newcommand{\shreyas}[1]{}
\definecolor{tableShade}{gray}{0.8}

\begin{document}
\title{The onset of hydrodynamically driven waving of marine grass in flow}
\author{Ravi Singh}
\affiliation{Brown University, Providence RI 02912 USA}
\author{L. Mahadevan}
\affiliation{Harvard University, Cambridge MA 02138 USA}
\author{M. M. Bandi}
\affiliation{Collective Interactions Unit, OIST Graduate University, 1919-1 Tancha, Onna-son, Okinawa, Japan 904-0495}
\author{Amala Mahadevan}
\affiliation{Woods Hole Oceanographic Institute, Woods Hole MA USA}
\author{Shreyas Mandre}
\affiliation{Brown University, Providence RI 02912 USA}

@@

We develop a self-consistent model of the flow past rigid vegetation, where the vegetation is represented by a continuum drag field. To explain the onset of spontaneous waving, we perform a linear stability analysis of the steady flow resulting from this model. We show that a threshold flow condition for the spontaneous waving of marine grass arises naturally in this framework due to the drag exerted by the vegetation on the flow. Beyond the threshold, we find two unstable modes; first is localized in a boundary layer near the canopy top, and second which spans the whole water column. The first mode shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it. The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations. 

@@

While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspects of existing theory remain unsatisfactory. 
(i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
(ii) Classical free shear flow is known to be unstable for all Reynolds numbers \cite{drazin}, whereas observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold 

@@

Here we present a self-consistent mathematical model that accounts correctly for these effects 

@@

The drag exerted by the grass bed on the flow is central to the hypothesized mechanism \cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
The flexibility of the grass blades, on the other hand, serves merely to aid in visualizing the flow structure resulting from the instability \cite{Nepf2012}. 
It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels \cite{Ghisal02}. 

@@

Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, 

@@

where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, and $N_g$ the grass number density per unit area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 

@@

where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 

@@

The Reynolds number (\Rey) of the flow based on the scale of a grass blade is expected to be $O(10^2-10^3)$, and hence the drag force is dominated by pressure and modeled as 

@@

normal to grass while $\bt,\bn$ are unit vectors along and normal to grass. 

@@

In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variations to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
Based on previous experiments \cite{Vivoni98}, we take $C_N \approx 1$.

@@

To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of eq.~\eqref{navier-stokes}, and examine its linear stability. 

@@

A comparison of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements is shown in fig.~\ref{basicflow}.
Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient. 

@@

\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 

@@

This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer \cite{Ghisal02} in the previous explanation of \monami. 

@@

This dependence of the boundary layer thickness on the grass density $N_g$, shown in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.

@@

To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)

@@

With these scalings, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 

@@

 parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are 

@@

of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of eq.~\eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 

@@

We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re($\sigma$)$>$0), emerges from the 
solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
The threshold Reynolds number increases with the grass density, indicating a competition between the 
destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of ad-hoc
assumptions for the mean velocity profile \cite{Raupach96,Delangre06}.

@@

Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in fig.~\ref{Re_vs_delta}b. 
The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
%Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.

\begin{table*}
\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
\renewcommand{\arraystretch}{1.4}
 \begin{tabular}{l|c|c|c}
			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
 Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
 Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
 Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
 Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
 Linearized dynamics	& $\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
 Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
 Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
 Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no
 \end{tabular}
 \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
 \label{tab:comparison}
\end{table*}
To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
The lowest $\Rey$ on the neutral curve sets the threshold. 
We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the most unstable mode to jump.

The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold. 
This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$. 
The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
\begin{align}
(D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \Rey \sigma (D^2-k^2)\phi,
\label{eqn:mode1asymp}
\end{align} 
in a region of thickness O($\delta$) near $y=\hg$.
Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see fig.~\ref{Re_vs_delta}). 
This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in fig.~\ref{Asymptotic_mode}, supporting this argument.

On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of eq.~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
This estimate simplifies eq.~\eqref{Orr-somerfield} to 
\begin{align}
\begin{split}
\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,  \quad &\text{ for $y<\hg$}  \\
\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, \quad &\text{ for $y>\hg$}.
\end{split}
\label{eqn:mode2asymp}
\end{align}
\begin{figure}
\includegraphics[]{Asymptotic_noshear}
\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. 
Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 1 shapes appear self-similar in shape as $\delta\to 0$, 
and are compared to each other on a rescaled coordinate in the inset. 
Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. 
The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
\label{Asymptotic_mode}
\end{figure}
Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in fig.~\ref{Asymptotic_mode}. 
We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role. The only remaining parameter $\Rey/\Ndg$ 
sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
%On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.

@@

The appearance of vegetation drag parameter in the dominant balances eq.~\eqref{eqn:mode1asymp} and eq.~\eqref{eqn:mode2asymp}, and the threshold criteria demonstrates its role in setting the threshold.

@@

Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami} \cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant \cite{Vivoni98,Ghisal02}. 
Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades. We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant. In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case. 

@@

The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy \cite{Ghisal02}. 

@@

\acknowledgments
MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan.

\bibliography{Grass}{}
\bibliographystyle{plain}
\end{document}
